Choosing climate forcing for numerical simulations of the Cordilleran ice sheet

### 2013-06-11 sections/intro.tex
Throw in all the modelling references
And expand the introduction.

>>>>

The numerical modelling of past glaciers and ice sheets allows indirect confrontation between glaciological theories and geomorphological reconstructions, challenging the validity of both. Yet a major obstacle in this exercise resides in the large uncertainties that affect climate forcing of the numerical ice sheet model \citep{hebeler-etal-2008}. This includes the uncertainty on Earth's present climate, and the larger uncertainty on its changes in the past.

Our area of focus for the present study is the Cordilleran Ice Sheet in north-western North America.
%In a numerical modelling perspective it is one of the least studied paleo-ice sheets on Earth, where nethertheless significant geomorphological data is available \citep{jackson-clague-1991}.
The Cordilleran Ice Sheet has been previously modelled as part of the North-american ice sheet complex \citep{marshall-clarke-1999,calov-etal-2002,tarasov-peltier-2004,bintanja-etal-2005,gregoire-etal-2012}, and within a larger-scale Arctic or Global numerical domain \citep{huybrechts-tsiobbel-1996,charbit-etal-2002,johnson-fastook-2002,zweck-huybrechts-2005,abeouchi-etal-2007,charbit-etal-2013}. In those studies, various methods have been used to provide climate forcing to ice-sheet models, ranging from simplistic parametrizations to coupling to General Circulation Models (GCM).

\begin{itemize}

  \item \citet{johnson-fastook-2002} used a parametrized climate forcing based on elevation and distance to a climatic north pole \citep{fastook-prentice-1994}.

  \item \citet{bintanja-etal-2005} used present-day reanalysis data as a reference climate to which they applied time-dependant temperature offsets to recreate glacial conditions.

  \item \citet{huybrechts-tsiobbel-1996} use palæo-climate simulation output from GCM to force an ice-sheet model at steady-state. Such GCMs are typically too computationally demanding to simulate climate evolution at the time-scales of ice sheet growth and decay. Therefore a common method to produce time-dependant climate forcing consist in interpolating in time between different GCM runs, either linearly \citep{charbit-etal-2002} or more commonly using a wieghting function based on ice-core record data \citep{marshall-clarke-1999,tarasov-peltier-2004,zweck-huybrechts-2005,gregoire-etal-2012}. GCM palæo-climate simulations typically rely on ice-sheet reconstructions of the type of ICE-4G for their surface topography boundary condition. In turn, ice-sheet models based on these simulations generally reproduce LGM ice extents closely consistent with the reconstruction.

  \item Finally \citet{yoshimori-etal-2001,calov-etal-2002,abeouchi-etal-2007,charbit-etal-2013} couple their ice-sheet models to GCM of coarser resolution and reduced complexity.

\end{itemize}

%TODO tarasov-peltier-1997

If these studies reproduce well the magnitude of the North American glaciation at Last Glacial Maximum, there is a general tendency for studies that do not undirectly rely on reconstructions of the type of ICE-4G to predict excessive ice cover in northern Yukon Territory and Alaska, regions that have not been covered by ice in hundreds of thousands of years \citep{johnson-fastook-2002,bintanja-etal-2005,huybrechts-tsiobbel-1996,charbit-etal-2002,marshall-clarke-1999,tarasov-peltier-2004,zweck-huybrechts-2005,gregoire-etal-2012}.

Here we use a Parallel Ice Sheet Model (PISM) \citep{web:pism} to simulate the entire Cordilleran Ice Sheet at the Last Glacial Maximum (LGM). We force our model with multiple climate datasets and compare our results to the mapped LGM ice sheet margin from \citet{kleman-etal-2010}. See \citet{quiquet-etal-2012} for a similar comparison on Greenland. To avoid circular dependance on prescribed ice sheet geometries, and allow high spatial resolution in our climate forcing, we use present-day climate data and given temperature offsets in a manner similar to \citet{bintanja-etal-2005}. To our knowledge, this is the first modelling study which specifically focus on the Cordilleran Ice Sheet since the one by \citep{robert-1991}, who focused specifically on its southern margin.

### 2013-06-11 sections/model.tex
Throw in all the modelling references
And expand the introduction.

Monthly temperature and precipitation are converted to surface accumulation and ablation by a temperature-index (positive degree-day) model. Accumulation is computed as the amount of precipitation that falls when temperature is below freezing\ask{Is this true? What is the current threshold in stable 0.5?}. Ablation is computed from the number of positive degree-days, defined as the integral of temperatures above 0\,\degree\,C in one year. Positive degree-days are calculated by the \citet{calov-greve-2011} formula to accounts for day-to-day temperature variability by assuming a normal distribution of variance $\sigma=5\degree C$ along a central (input) value $T$:

>>>>

Monthly temperature and precipitation are converted to surface accumulation and ablation by a temperature-index (positive degree-day) model. Accumulation is computed as the amount of precipitation that falls when temperature is below freezing\ask{Is this true? What is the current threshold in stable 0.5?}. Ablation is computed from the number of positive degree-days, defined as the integral of temperatures above 0\,\degree\,C in one year. Positive degree-days are calculated by the \citet{calov-greve-2005} formula to accounts for day-to-day temperature variability by assuming a normal distribution of variance $\sigma=5\degree C$ along a central (input) value $T$:
